This element is similar to {\tt TRFMODE}, but it allows faster simulation of more than
one mode.  Also, the mode data is specified in an SDDS file.  This file can be
generated using the APS version of URMEL, or by hand. It must have the following
columns and units:
\begin{enumerate}
\item {\tt Frequency} --- The frequency of the mode in Hz.  Floating point.
\item {\tt Q} --- The quality factor.  Floating point.
\item {\tt ShuntImpedance} or {\tt ShuntImpedanceSymm} --- The shunt
  impedance in Ohms/m, defined as $V^2/(2*P)/x$ or $V^2/(2*P)/y$.
  Floating point. By default, {\tt ShuntImpedance} is used.  However,
  if the parameter \verb|USE_SYMM_DATA| is non-zero, then {\tt
  ShuntImpedanceSymm} is used.  The latter is the full-cavity shunt
  impedance that URMEL computes by assuming that the input cavity used
  is one half of a symmetric cavity.
\end{enumerate}

The file may also have the following columns:
\begin{enumerate}
\item {\tt beta} --- Normalized load impedance (dimensionless).  Floating point. If not given, the
 $\beta=0$ is assumed for all modes. 
\item {\tt xMode} --- If given, then only modes for which the value is nonzero will produce an
 x-plane kick.  Integer.  If not given, all modes affect the x plane.
\item {\tt yMode} --- If given, then only modes for which the value is nonzero will produce an
 y-plane kick.  Integer.  If not given, all modes affect the y plane.
\end{enumerate}
    
{\bf N.B.:} Unlike {\tt TRFMODE}, {\tt FTRFMODE} does not include the longitudinal field that,
strictly speaking, must also be excited.  Generally this is a very small effect.  It will be
added in a future version.

In many simulations, a transient effect may occur when using this
element because, in the context of the simulation, the impedance is
switched on instantaneously.  This can give a false indication of the
threshold for instability. The {\tt RAMP\_PASSES} parameter should
be used to prevent this by slowly ramping the impedance to full
strength.  This idea is from M. Blaskiewicz (BNL).
