The input file for this element gives the transverse-wake Green
functions, $W_x(t)$ and $W_y(t)$, versus time behind the particle. The
units of the wakes are V/C/m, so this element simulates the integrated
wake of some structure (e.g., a cell or series of cells).  If you
have, for example, the wake for a cell and you need the wake for N
cells, then you may use the {\tt FACTOR} parameter to make the
appropriate multiplication.  The values of the time coordinate should
begin at 0 and be equi-spaced.  A positive value of time represents
the distance behind the exciting particle.   Time values must be equally
spaced.

The sign convention for $W_q$ ($q$ being $x$ or $y$) is as follows: a
particle with $q>0$ will impart a positive kick ($\Delta q^\prime >
0$) to a trailing particle following $t$ seconds behind if $W_q(t)>0$.
A physical wake function should be zero at $t=0$ and also be initially
positive as $t$ increases from 0.

Use of the {\tt CHARGE} parameter on the {\tt TRWAKE} element is
disparaged.  It is preferred to use the {\tt CHARGE} element as part
of your beamline to define the charge.  

Setting the {\tt N\_BINS} paramater to 0 is recommended.  This results
in auto-scaling of the number of bins to accomodate the beam.  The bin
size is fixed by the spacing of the time points in the wake.

The default degree of smoothing ({\tt SG\_HALFWIDTH=4}) may be excessive.
It is suggested that users vary this parameter to verify that results
are reliable if smoothing is employed ({\tt SMOOTHING=1}).

The {\tt XFACTOR} and {\tt YFACTOR} parameters can be used to adjust
the strength of the wakes if the location at which you place the {\tt
TRWAKE} element has different beta functions than the location at
which the object that causes the wake actually resides.  

The {\tt XPOWER} and {\tt YPOWER} parameters can be used to change the
dependence of the wake on the x and y coordinates, respectively, of
the particles.  Normally, {\tt XPOWER=1} and {\tt YPOWER=1}.  This is
an ordinary dipole wake in a (supposedly) symmetric chamber.  

If you have an asymmetric chamber, then you will have a transverse
wake kick even if the beam is centered.  (Of course, you'll need a 3-D
wake code like GdfidL or MAFIA to compute this wake.)  This part of
the transverse wake is described with {\tt XPOWER=0} and {\tt
YPOWER=0}.  It will result in an orbit distortion, but conceivably
could have other effects, such as emittance dilution.  If {\tt
XPOWER=0} or {\tt YPOWER=0}, the units for the x or y wake
(respectively) must be $V/C$.  A negative value of the wake
corresponds to a kick toward negative x (or y).
