This elements allows transformation of the twiss parameters of 
a beam with a first-order matrix.  The matrix is computed in 
various ways based on initial and final twiss parameters.  Depending on 
how you set it up, the final twiss parameters for your
beam may not be the twiss parameters you specify.  

The twiss parameter values \verb|BETAX|, \verb|BETAY|, etc. specified in the
element definition specify the target values of the transformation.
To completely specify the transformation, one must know the initial
values as well.  

{\bf Lattice-Function-Based Transformation}

If \verb|FROM_BEAM| is zero, which is the default, then the initial
values are taken from the incoming lattice functions computed by
\verb|twiss_output|.  This provides a way to transform the lattice
functions between two parts of a transport line without designing
intervening optics.  A beam that is matched at the beginning of the
transport line will remain matched.  A beam that is mismatched at the
beginning of the transport line {\em will not} be matched after
the \verb|TWISS| element.

By default, each time the twiss parameters are recomputed, the
transformation is updated to maintain the desired lattice functions at
the exit of the \verb|TWISS| element. Setting \verb|COMPUTE_ONCE| to a
non-zero value specifies that {\tt elegant} should compute the
transformation matrix only once, i.e., for the first set of computed
lattice functions.

By default, the transformation is applied to the beam only the first
time it passes the element.  Setting \verb|APPLY_ONCE| to a zero
will result in application of the transformation at each pass.

{\bf Beam-Ellipse-Based Transformation}

If \verb|FROM_BEAM| is non-zero, the the initial values for the transformation
are computed from a beam.  This provides a way to transform the beam ellipse
to the desired twiss parameters irrespective of the lattice.  The results
from \verb|twiss_output| will not necessarily be matched downstream of this
element.  Only if the beam ellipse and lattice ellipse are the same will this
occur.

By default, each time a new beam is generated, the transformation will
be updated to maintain the desired beam ellipse at the exit of the
\verb|TWISS| element.  Setting \verb|COMPUTE_ONCE| to a non-zero value
specifies that {\tt elegant} should compute the transformation matrix
only once, i.e., for the first beam it sees.

By default, the transformation is applied to the beam only the first
time it passes the element.  Setting \verb|APPLY_ONCE| to a zero
will result in application of the transformation at each pass.  This
would make sense, for example, if the \verb|TWISS| element was
filling in for a section of a ring.  It wouldn't make sense if
the \verb|TWISS| element was being used to match the beam from
a transport line to a ring.

