This element simulates a sextupole using a kick method based on
symplectic integration.  The user specifies the number of kicks and
the order of the integration.  For computation of twiss parameters,
chromaticities, and response matrices, this element is treated like a
standard thick-lens sextuupole; i.e., the number of kicks and the
integration order become irrelevant.

\begin{raggedright}
Specification of systematic and random multipole errors is supported
through the \verb|SYSTEMATIC_MULTIPOLES| and 
\verb|RANDOM_MULTIPOLES|
fields.  These fields give the names of SDDS files that supply the
multipole data.  The files are expected to contain a single page of
data with the following elements:
\end{raggedright}
\begin{enumerate}
\item Floating point parameter {\tt referenceRadius} giving the reference
 radius for the multipole data.
\item An integer column named {\tt order} giving the order of the multipole.
The order is defined as $(N_{poles}-2)/2$, so a quadrupole has order 1, a
sextupole has order 2, and so on.
\item Floating point columns {\tt an} and {\tt bn} giving the values for the
normal and skew multipole strengths, respectively.  These are defined as a fraction 
of the main field strength measured at the reference radius, R: 
$a_n  = \frac{K_n r^n / n!}{K_m r^m / m!}$, where 
$m=2$ is the order of the main field and $n$ is the order of the error multipole.
A similar relationship holds for the skew multipoles.
For random multipoles, the values are interpreted as rms values for the distribution.
\end{enumerate}

Apertures specified via an upstream \verb|MAXAMP| element or an \verb|aperture_input|
command will be imposed inside this element, with the following rules/limitations:
\begin{itemize}
\item If apertures from both sources are present, the smallest is used.
\item The apertures are assumed to be rectangular, even if the \verb|ELLIPTICAL| qualifier
 is set for \verb|MAXAMP|.
\end{itemize}
