This element is used to generate multiple dimension (1, 2, 4, or 6)
histogram output of particle coordinates.  

The calculation is set up through output filename: \verb|FILE1D|,
\verb|FILE2DH|, \verb|FILE2DV|, \verb|FILE2DL|, \verb|FILE4D|, \verb|FILE6D|.  
They may be an incomplete filename (see \verb|HISTOGRAM| for
detail). If \verb|LUMPED| set to non zero, then results are directed
to a multi page SDDS file with each page contains data of same
elements \verb|MHISTOGRAM| but at difference occurrence instead of
multiple SDDS files. In this case the ``\%ld'' in filename is ignored.

The bin number used to do histogram analysis is given through a SDDS
file from \verb|INPUT_BINS|. It contains 4 columns: \verb|Bins_1D|,
\verb|Bins_2D|, \verb|Bins_4D|, \verb|Bins_6D|; and 6 rows (x, x', y,
y', dt, delta). A non-zero value in \verb|Bins_1D| is a switch for
doing histogram analysis in corresponding dimension, and the maximum
value in \verb|Bins_1D| is used as bin number to do the analysis.

The normalization is different from \verb|HISTOGRAM| as we alwayse
treat bin-size = 1.


\vspace{5mm}
The output file uses the general format designed for a 
n-dimensional histogram data. It must contains a column named
``Frequency'' (Type: ``double''), and following parameters:
\begin{itemize}
\item \verb|ND| --- Type: long; Value: ``n''.
\item \verb|Variable??Name| --- Type: ``string''. ``??'' counts from ``0''
to ``ND-1'' in double digits format, same for all following parameters.
\item  \verb|Variable??Min| --- Type: ``double''. Minimum value of ``??'' variable.
\item  \verb|Variable??Max| --- Type: ``double''. Maximum value of ``??'' variable.
\item  \verb|Variable??Interval| --- Type: ``double''. Bin size of ``??'' variable.
\item  \verb|Variable??Dimension| --- Type: ``long''. Total number of bins of ``??'' variable.
     Variable??Dimension = (Variable??Max - Variable??Min)/Variable??Interval+1.
\end{itemize} 
The data is arranged as it has a ``ND'' index counter $[i_{ND-1} |
... | i_1]$, where $i_{ND-1}$ takes value from ``0'' to
``Variable[\%02d ND-1]Dimension''.
 