This element simulates a quadrupole using a matrix of first, second,
or third order.

By default, the element has hard edges and constant field within the
defined length, {\tt L}.  However, this element supports two different methods of implementing fringe fields.
Which method is used is determined by the \verb|FRINGE_TYPE| parameter.

The first method is based on a third-order matrix formalism and the assumption of
linear fringe fields.  To invoke this method, one specifies ``inset'' or
``fixed-strength'' for the \verb|FRINGE_TYPE| parameter and then provides
a non-zero value for {\tt FFRINGE}. If  {\tt FFRINGE} is zero (the default), then the magnet
is hard-edged regardless of the setting of \verb|FRINGE_TYPE|.  If {\tt FFRINGE} is positive, then the magnet has
linear fringe fields of length {\tt FFRINGE*L/2} at each end.  That
is, the total length of fringe field from both ends combined is {\tt
FFRINGE*L}.

Depending on the value of {\tt FRINGE\_TYPE}, the fringe fields are
modeled as contained within the length {\tt L} (``inset'' type) or
extending symmetrically outside the length {\tt L} (``fixed-strength''
type).

For ``inset'' type fringe fields, the length of the ``hard core'' part of
the quadrupole is {\tt L*(1-FFRINGE)}.  For ``fixed-strength'' type fringe fields,
the length of the hard core is {\tt L*(1-FFRINGE/2)}.  In the latter case,
the fringe gradient reaches 50\% of the hard core value at the nominal boundaries
of the magnet. This means that the integrated strength of the magnet does not
change as the {\tt FFRINGE} parameter is varied. This is not the case with
``inset'' type fringe fields.

A more recent implementation of fringe field effects is based on integrals and is invoked
by setting \verb|FRINGE_TYPE| to ``integrals''.  However, this method provides a first-order matrix only.
This is based on publications of D.  Zhuo {\em et al.} \cite{Zhou-IPAC10} and  J. Irwin {\em et
  al.} \cite{Irwin-PAC95}, as well as unpublished work of C. X. Wang (ANL).  The fringe field is characterized by 
10 integrals given in equations 19, 20, and 21 of \cite{Zhou-IPAC10}.  However, the values input into {\tt elegant}
should be normalized by $K_1$ or $K_1^2$, as appropriate.

For the exit-side fringe field, let $s_1$ be the center of the magnet, $s_0$ be the location of the nominal end of the magnet
(for a hard-edge model), and let $s_2$ be a point well outside the magnet.  
Using $K_{1,he}(s)$ to represent the hard edge model and $K_1(s)$ the actual field profile, we 
define the normalized difference as $\tilde{k}(s) = (K_1(s) - K_{1,he}(s))/K_1(s_1)$.  (Thus, $\tilde{k}(s) = \tilde{K}(s)/K_0$, using
the notation of Zhou {\em et al.})

The integrals to be input to {\tt elegant} are defined as 
\begin{eqnarray}
i_0^- = \int_{s_1}^{s_0} \tilde{k}(s) ds & & i_0^+ = \int_{s_0}^{s_2} \tilde{k}(s) ds \\
i_1^- = \int_{s_1}^{s_0} \tilde{k}(s) (s-s_0) ds & & i_1^+ = \int_{s_0}^{s_2} \tilde{k}(s) (s-s_0) ds \\
i_2^- = \int_{s_1}^{s_0} \tilde{k}(s) (s-s_0)^2 ds & & i_2^+ = \int_{s_0}^{s_2} \tilde{k}(s) (s-s_0)^2 ds \\
i_3^- = \int_{s_1}^{s_0} \tilde{k}(s) (s-s_0)^3 ds & & i_3^+ = \int_{s_0}^{s_2} \tilde{k}(s) (s-s_0)^3 ds \\
\lambda_2^- = \int_{s_1}^{s_0} ds \int_s^{s_0} ds\prime \tilde{k}(s) \tilde{k}(s\prime) (s\prime-s) & & 
\lambda_2^+ = \int_{s_0}^{s_2} ds \int_s^{s_2} ds\prime \tilde{k}(s) \tilde{k}(s\prime) (s\prime-s) 
\end{eqnarray}

Normally, the effects are dominated by $i_1^-$ and $i_1^+$.  
